Self Referencing Sequences

### 2014-11-06 draft.tex
Adding New Figures
Edit Single Typo in Determinism
Otherwise Determinism is Done

We will study a sequences which are self-referencing. That is, each number in the sequence will correspond to a previous number in the same sequence. In particular, we will analyze self-referencing sequences with respect to block lengths. Consider the sequence
\begin{equation}
\label{eq:onetwo}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{equation}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
\begin{figure}
\label{fig:sequenceblocks}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks.pdf}
\caption{Sequence broken up into contiguous blocks with lengths generating same sequence.}
\end{figure}

@@

With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is shown in Figure~\ref{fig:onetwothree}.
\begin{figure}
\label{fig:onetwothree}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
\caption{Sequence generated by $(123)$ broken up into contiguous blocks.}
\end{figure}

@@

We claim that a self-referencing sequence is uniquely determined by the generating cycle $C$ and the starting number $a \in C$. We will prove this by showing that if we are given the first number of the sequence, there is only one way to extend the sequence. In order to do this, we formalize the idea of breaking up a sequence into blocks.

@@

\begin{proposition}
A self-referencing sequence $\{a_i\}$ is uniquely specified by the generating cycle $C$ and the starting number $a \in C$. In other words, such an $\{a_i\}$ is guaranteed to exist and is unique.
\end{proposition}
\begin{proof} 
The proof is constructive. The self-referential property of the sequence implies that the contents of block $B_i$ are entirely determined by the values of $a$, $i$, and $s_i$. The value of $i$ determines which number is repeated throughout the block. For example, when $i = 1$, the number is the starting number $a$, and when $i = 2$, we use the number following the starting number in $C$, and so on for larger values of $i$. The value of $s_i$ simply dictates the length of block $B_i$.

Thus, the sequence is uniquely given by first ``writing down" $B_1$, followed by $B_2$, followed by $B_3$, etc. One detail is that it remains to check that $s_i$ is known when $B_i$ is written down. However, writing down a block always involves writing down at least one element of the sequence, so the value of $s_i$ is always known when $B_i$ is written down. Note that there is an edge case where $s_i$ is not yet written down if the starting number is 1 and $i =2$. However, in this case, the value of $s_i$ is still known (as it is given by the rule) and thus $B_i$ can still be written down.
\end{proof}
%
%\begin{lemma}
%Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
%\end{lemma}
%
%\begin{proof}
%Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
%
%\begin{figure}{hb}
%\label{fig:unequalsequences}
%\center
%\includegraphics[width=0.5\linewidth]{lemma1.pdf}
%\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
%\end{figure}
%
%This implies that every block before the $k$th block just contains 1's since each block refered it itself. This means that either $j = 2$, or the sequence is not self-referential. But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle. 
%\end{proof}
%
%
%
%We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element. Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$. For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
%\begin{quote}
%\textbf{Algorithm:}\\
%Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
%
%This procedure continues deterministically so long as there is an element to read. Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
%\end{quote}

>>>>

We will study a sequences which are self-referencing. That is, each number in the sequence will correspond to a previous number in the same sequence. In particular, we will analyze self-referencing sequences with respect to block lengths. Consider the sequence
\begin{equation}
\label{eq:onetwo}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{equation}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
\begin{figure}
\label{fig:sequenceblocks}
\center
\includegraphics[width=0.4\linewidth]{sequenceBlocks}
\caption{Sequence broken up into contiguous blocks with lengths generating same sequence.}
\end{figure}

@@

With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is shown in Figure~\ref{fig:onetwothree}.
\begin{figure}
\label{fig:onetwothree}
\center
\includegraphics[width=0.4\linewidth]{sequenceBlocks123}
\caption{Sequence generated by $(123)$ broken up into contiguous blocks.}
\end{figure}

@@

We claim that a self-referencing sequence is uniquely determined by the generating cycle $C$ and the starting number $c \in C$. We will prove this by showing that if we are given the first number of the sequence, there is only one way to extend the sequence. In order to do this, we formalize the idea of breaking up a sequence into blocks.

@@

\begin{proposition}
A self-referencing sequence $\{a_i\}$ is uniquely specified by the generating cycle $C$ and the starting number $c \in C$. In other words, such an $\{a_i\}$ is guaranteed to exist and is unique.
\end{proposition}
\begin{proof} 
The proof is constructive. The self-referential property of the sequence implies that the contents of block $B_i$ are entirely determined by the values of $c$, $i$, and $a_i$. The value of $i$ determines which number is repeated throughout the block. For example, when $i = 1$, the number is the starting number $c$, and when $i = 2$, we use the number following the starting number in $C$, and so on for larger values of $i$. The value of $a_i$ simply dictates the length of block $B_i$.

Thus, the sequence is uniquely given by first ``writing down" $B_1$, followed by $B_2$, followed by $B_3$, etc. One detail is that it remains to check that $a_i$ is known when $B_i$ is written down. However, writing down a block always involves writing down at least one element of the sequence, so the value of $a_i$ is always known when $B_i$ is written down. Note that there is an edge case where $a_i$ is not yet written down if the starting number is 1 and $i =2$. However, in this case, the value of $a_i$ is still known (as it is given by the rule) and thus $B_i$ can still be written down.
\end{proof}
%
%\begin{lemma}
%Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
%\end{lemma}
%
%\begin{proof}
%Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
%
%\begin{figure}{hb}
%\label{fig:unequalsequences}
%\center
%\includegraphics[width=0.5\linewidth]{lemma1.pdf}
%\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
%\end{figure}
%
%This implies that every block before the $k$th block just contains 1's since each block refered it itself. This means that either $j = 2$, or the sequence is not self-referential. But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle. 
%\end{proof}
%
%
%
%We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element. Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$. For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
%\begin{quote}
%\textbf{Algorithm:}\\
%Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
%
%This procedure continues deterministically so long as there is an element to read. Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
%\end{quote}
